BIP-Predictor

### 2025-05-26 Latex/BIP_predictor.tex
up

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:26 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-07-01 & 902.571 & 90.700 & 85.800 & 92.500 & -1.300 \\ 
  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
  2024-09-01 &  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
  2024-10-01 & 900.761 & 90.800 & 84.100 & 90.700 & -1.200 \\ 
  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-12-01 &  &  & 82.800 & 86.900 & -0.800 \\ 
  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Ragged end.  } 
\label{f_stat}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:26 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-05-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
  2024-06-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
  2024-07-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
  2024-08-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-09-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
  2024-10-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Data aligned at sample end.} 
\label{f_stat}
\end{table}

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:28 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h= 0 & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
  \hline
HP-C filtered & 0.000 & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
  Unfiltered & 0.028 & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
   \hline
\end{tabular}
\caption{Statistical significance of unfiltered and HP-C filtered forecasts based on HAC-adjusted Wald test: all indicators and full sample (without pandemic).  } 
\label{f_stat}
\end{table}
We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.

@@

Figure \ref{bk_gammak} displays the coefficients of the multivariate nowcasts, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags. The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons. The M-SSA nowcast adheres to the specified constraint, exhibiting approximately one-third fewer zero-crossings than the M-MSE, confirming the effectiveness of the smoothness control imposed during the design. This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations. Increased smoothness achieved by the M-SSA comes at the cost of a slight reduction in tracking accuracy, as indicated by a marginally lower target correlation. This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/mssa_msse_now.pdf}\caption{Two-sided HP(160) applied to GDP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green). The dashed vertical line delimits in- and out-of-sample spans. The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:28 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & M-SSA & M-MSE \\ 
  \hline
Target correlation & 0.68 & 0.69 \\ 
  HT & 6.50 & 3.71 \\ 
   \hline
\end{tabular}
\caption{Sample target correlations and HTs of M-SSA and M-MSE nowcasts.  } 
\label{corhtnow}
\end{table}The results validate that the smoothing constraint incorporated into the multivariate extension of Criterion \eqref{critssa} effectively manages the zero-crossings of the predictor: sample estimates of the holding times (HTs) align with the expected values derived from Equation \eqref{ht}\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package. This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.}. An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this tradeoff could be beneficial in our application. Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.

@@

There are various combinations of M-SSA components for tracking BIP. For illustration, we here select the single BIP component resulting from the M-SSA filter tracking HP-BIP. Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample. 
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:32 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of BIP (M-SSA) component optimized for forecast horizons $h=0,...,6$  (the columns) on HP-BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_hpbip_wc}
\end{table}Our choice—focusing on the single M-SSA BIP component—emphasizes simplicity and interpretability, as future BIP and HP-BIP are connected through their shared low-frequency component. Additionally, the M-SSA BIP component is the most important predictor in the regression of the M-SSA outputs on BIP.  Finally, revisions due to quarterly updates of the regression equations are minimized by this predictor, and the estimates are more stable and robust over time\footnote{Other combinations can be experimented with using the M-SSA package. Regression weights of more complex designs, involving multiple components, tend to be more difficult to interpret due to the occurrence of negative weights and potential sign changes of the regressors over time, which may indicate overfitting or non-stationarity.} (see Section \eqref{revisions}). Note that information from the other  indicators is incorporated into the M-SSA BIP component through the multivariate filter design.\\

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:33 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.086 & 0.049 & 0.018 & 0.005 & 0.002 & 0.002 & 0.042 \\ 
  Shift=1 & 0.238 & 0.165 & 0.097 & 0.033 & 0.005 & 0.001 & 0.005 \\ 
  Shift=2 & 0.815 & 0.680 & 0.428 & 0.180 & 0.029 & 0.007 & 0.008 \\ 
  Shift=3 & 0.866 & 0.963 & 0.922 & 0.699 & 0.269 & 0.079 & 0.047 \\ 
  Shift=4 & 0.361 & 0.808 & 0.989 & 0.954 & 0.546 & 0.127 & 0.034 \\ 
  Shift=5 & 0.351 & 0.786 & 0.982 & 0.943 & 0.579 & 0.245 & 0.042 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:33 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.950 & 0.917 & 0.877 & 0.851 & 0.865 & 0.910 & 0.960 \\ 
  Shift=1 & 1.028 & 1.007 & 0.960 & 0.899 & 0.867 & 0.878 & 0.912 \\ 
  Shift=2 & 1.040 & 1.043 & 1.030 & 0.984 & 0.933 & 0.910 & 0.913 \\ 
  Shift=3 & 1.006 & 1.024 & 1.040 & 1.025 & 0.983 & 0.952 & 0.941 \\ 
  Shift=4 & 0.977 & 1.003 & 1.034 & 1.040 & 1.005 & 0.964 & 0.937 \\ 
  Shift=5 & 0.984 & 1.011 & 1.044 & 1.057 & 1.031 & 0.986 & 0.949 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean_wc}
\end{table}

@@

\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs. false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).\label{ROC_GDP_shift_1_4}}\end{center}\end{figure}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:36 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
  \hline
Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
   \hline
\end{tabular}
\caption{AUCs of predictors against GDP at various forecast horizons.} 
\label{p_val_wc5}
\end{table}At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters. However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA. These results confirm that controlling the rate of sign changes of the predictor, while maximizing its target correlation through M-SSA can enhance predictor performance in terms of sign accuracy and false alarm rate.

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:36 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.086 & 0.049 & 0.018 & 0.005 & 0.002 & 0.002 & 0.042 \\ 
  Shift=1 & 0.238 & 0.165 & 0.097 & 0.033 & 0.005 & 0.001 & 0.005 \\ 
  Shift=2 & 0.815 & 0.680 & 0.428 & 0.180 & 0.029 & 0.007 & 0.008 \\ 
  Shift=3 & 0.866 & 0.963 & 0.922 & 0.699 & 0.269 & 0.079 & 0.047 \\ 
  Shift=4 & 0.361 & 0.808 & 0.989 & 0.954 & 0.546 & 0.127 & 0.034 \\ 
  Shift=5 & 0.351 & 0.786 & 0.982 & 0.943 & 0.579 & 0.245 & 0.042 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc5}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:36 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.950 & 0.917 & 0.877 & 0.851 & 0.865 & 0.910 & 0.960 \\ 
  Shift=1 & 1.028 & 1.007 & 0.960 & 0.899 & 0.867 & 0.878 & 0.912 \\ 
  Shift=2 & 1.040 & 1.043 & 1.030 & 0.984 & 0.933 & 0.910 & 0.913 \\ 
  Shift=3 & 1.006 & 1.024 & 1.040 & 1.025 & 0.983 & 0.952 & 0.941 \\ 
  Shift=4 & 0.977 & 1.003 & 1.034 & 1.040 & 1.005 & 0.964 & 0.937 \\ 
  Shift=5 & 0.984 & 1.011 & 1.044 & 1.057 & 1.031 & 0.986 & 0.949 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean6}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:36 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.104 & 1.066 & 1.019 & 0.989 & 1.006 & 1.058 & 1.116 \\ 
  Shift=1 & 1.232 & 1.206 & 1.150 & 1.076 & 1.038 & 1.051 & 1.092 \\ 
  Shift=2 & 1.146 & 1.150 & 1.136 & 1.085 & 1.028 & 1.002 & 1.006 \\ 
  Shift=3 & 1.008 & 1.027 & 1.042 & 1.028 & 0.986 & 0.954 & 0.943 \\ 
  Shift=4 & 0.988 & 1.014 & 1.045 & 1.051 & 1.016 & 0.975 & 0.948 \\ 
  Shift=5 & 0.967 & 0.995 & 1.027 & 1.039 & 1.013 & 0.970 & 0.933 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_direct_without_covid7}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:24:36 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
  \hline
RRMSE & 0.860 & 0.835 & 0.907 & 0.998 & 0.989 & 1.017 \\ 
   \hline
\end{tabular}
\caption{rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_direct_mean_without_covid8}
\end{table}

>>>>

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:26:22 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-07-01 & 902.571 & 90.700 & 85.800 & 92.500 & -1.300 \\ 
  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
  2024-09-01 &  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
  2024-10-01 & 900.761 & 90.800 & 84.100 & 90.700 & -1.200 \\ 
  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-12-01 &  &  & 82.800 & 86.900 & -0.800 \\ 
  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Ragged end.  } 
\label{f_stat}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:26:22 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-05-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
  2024-06-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
  2024-07-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
  2024-08-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-09-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
  2024-10-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Data aligned at sample end.} 
\label{f_stat}
\end{table}

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:26:23 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h= 0 & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
  \hline
HP-C filtered & 0.000 & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
  Unfiltered & 0.028 & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
   \hline
\end{tabular}
\caption{Statistical significance of unfiltered and HP-C filtered forecasts based on HAC-adjusted Wald test: all indicators and full sample (without pandemic).  } 
\label{f_stat}
\end{table}
We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.

@@

Figure \ref{bk_gammak} displays the coefficients of the multivariate nowcasts, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags. The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons. The M-SSA nowcast adheres to the specified constraint, exhibiting approximately one-third fewer zero-crossings than the M-MSE, confirming the effectiveness of the smoothness control imposed during the design. This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations. Increased smoothness achieved by the M-SSA comes at the cost of a slight reduction in tracking accuracy, as indicated by a marginally lower target correlation. This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/mssa_msse_now.pdf}\caption{Two-sided HP(160) applied to GDP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green). The dashed vertical line delimits in- and out-of-sample spans. The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:26:35 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & M-SSA & M-MSE \\ 
  \hline
Target correlation & 0.68 & 0.69 \\ 
  HT & 6.50 & 3.71 \\ 
   \hline
\end{tabular}
\caption{Sample target correlations and HTs of M-SSA and M-MSE nowcasts.  } 
\label{corhtnow}
\end{table}The results validate that the smoothing constraint incorporated into the multivariate extension of Criterion \eqref{critssa} effectively manages the zero-crossings of the predictor: sample estimates of the holding times (HTs) align with the expected values derived from Equation \eqref{ht}\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package. This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.}. An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this tradeoff could be beneficial in our application. Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.

@@

There are various combinations of M-SSA components for tracking BIP. For illustration, we here select the single BIP component resulting from the M-SSA filter tracking HP-BIP. Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample. 
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:27:26 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of BIP (M-SSA) component optimized for forecast horizons $h=0,...,6$  (the columns) on HP-BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_hpbip_wc}
\end{table}Our choice—focusing on the single M-SSA BIP component—emphasizes simplicity and interpretability, as future BIP and HP-BIP are connected through their shared low-frequency component. Additionally, the M-SSA BIP component is the most important predictor in the regression of the M-SSA outputs on BIP.  Finally, revisions due to quarterly updates of the regression equations are minimized by this predictor, and the estimates are more stable and robust over time\footnote{Other combinations can be experimented with using the M-SSA package. Regression weights of more complex designs, involving multiple components, tend to be more difficult to interpret due to the occurrence of negative weights and potential sign changes of the regressors over time, which may indicate overfitting or non-stationarity.} (see Section \eqref{revisions}). Note that information from the other  indicators is incorporated into the M-SSA BIP component through the multivariate filter design.\\

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:27:36 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:27:36 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean_wc}
\end{table}

@@

\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs. false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).\label{ROC_GDP_shift_1_4}}\end{center}\end{figure}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:28:11 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
  \hline
Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
   \hline
\end{tabular}
\caption{AUCs of predictors against GDP at various forecast horizons.} 
\label{p_val_wc5}
\end{table}At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters. However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA. These results confirm that controlling the rate of sign changes of the predictor, while maximizing its target correlation through M-SSA can enhance predictor performance in terms of sign accuracy and false alarm rate.

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:28:11 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc5}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:28:11 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean6}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:28:11 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & 0.949 & 0.934 \\ 
  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & 0.941 \\ 
  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & 0.936 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_direct_without_covid7}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Mon May 26 07:28:11 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
  \hline
RRMSE & 0.862 & 0.838 & 0.912 & 1.008 & 1.001 & 1.028 \\ 
   \hline
\end{tabular}
\caption{rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_direct_mean_without_covid8}
\end{table}
